{\Huge Supersymmetry} \\ \vspace{5mm} Lecture notes for FYS5190/FYS9190 Supersymmetry

### 2022-01-01 groups.tex
Typos Kevin

\df{A {\bf Lie group} $G$ is a finite-dimensional {\bf smooth manifold} where group multiplication and inversion are smooth functions, meaning that given elements $g(\mathbf{a}), g(\mathbf{a}') \in G$, $g(\mathbf{a}')\circ g(\mathbf{a}') = g(\mathbf{b})$ where $\mathbf{b}(\mathbf{a}, \mathbf{a}')$ is a smooth function of $\mathbf a$ and $\mathbf a'$, and $g^{-1}(\mathbf{a}) = g(\mathbf{b})$ where $\mathbf{b}(\mathbf{a})$ is a smooth function of $\mathbf a$.}

@@

From our earlier example we immediately see that the translation group $T(1)$, given the parameterisation of the elements $g(a) = e^{a\frac{d}{dx}}$, is a Lie group since $g(a)\cdot g(a') =g(b)= g(a+a')$ and $b=a+a'$ is an analytic function of $a$ and $a'$, and for the inverse $g^{-1}(a)=g(b)=g(-a)$ where $b=-a$ is a smooth function of $a$. Here, we can write the action of the group on the vector space $\mathbb{R}^1$ as $x'=f(x,a) = x+a$. The matrix groups groups are also Lie groups as they have $n\times n$-matrix representations where the matrices $A$ are parametrised by a certain number of parameters $\mathbf a$,  so that $x_i'=f_i(\mathbf{x}, \mathbf a) = [A(\mathbf a)\mathbf{x}]_i$.

>>>>

\df{A {\bf Lie group} $G$ is a finite-dimensional {\bf smooth manifold} where group multiplication and inversion are smooth functions, meaning that given elements $g(\mathbf{a}), g(\mathbf{a}') \in G$, $g(\mathbf{a})\circ g(\mathbf{a}') = g(\mathbf{b})$ where $\mathbf{b}(\mathbf{a}, \mathbf{a}')$ is a smooth function of $\mathbf a$ and $\mathbf a'$, and $g^{-1}(\mathbf{a}) = g(\mathbf{b})$ where $\mathbf{b}(\mathbf{a})$ is a smooth function of $\mathbf a$.}

@@

From our earlier example we immediately see that the translation group $T(1)$, given the parameterisation of the elements $g(a) = e^{a\frac{d}{dx}}$, is a Lie group since $g(a)\cdot g(a') =g(b)= g(a+a')$ and $b=a+a'$ is an analytic function of $a$ and $a'$, and for the inverse $g^{-1}(a)=g(b)=g(-a)$ where $b=-a$ is a smooth function of $a$. Here, we can write the action of the group on the vector space $\mathbb{R}^1$ as $x'=f(x,a) = x+a$. The matrix groups are also Lie groups as they have $n\times n$-matrix representations where the matrices $A$ are parametrised by a certain number of parameters $\mathbf a$,  so that $x_i'=f_i(\mathbf{x}, \mathbf a) = [A(\mathbf a)\mathbf{x}]_i$.

### 2022-01-01 poincare.tex
Typos Kevin

So, what is the second Casimir of the Poincaré algebra?
\df{We define the {\bf Pauli-Ljubanski polarisation vector} by:
\begin{equation}
W_\mu \equiv \frac{1}{2} \epsilon_\mu{}_ \nu{}_\rho{}_\sigma P^\nu M^{\rho\sigma}.
\end{equation}
where $ \epsilon_\mu{}_ \nu{}_\rho{}_\sigma$ is the totally antisymmetric Levi-Civita tensor with $ \epsilon_{0123}=1$.}
We can show that this vector is translation invariant, {\it i.e.}\ that it commutes with the translation operator,
\begin{equation}
[P_\mu, W_\nu]=0.
\label{eq:PW_commutator}
\end{equation}
Then $W^2 = W_\mu W^\mu$ is a Casimir operator of the Poincaré algebra since we can show that
\begin{eqnarray}
\left[P_\mu, W^2\right] &=& 0.\\
\left[M_\mu{}_\nu, W^2\right] &=& 0,
\end{eqnarray}
Note that the second of these relationships is not trivial to demonstrate. See \cite{IntrSUSY2010} for a complete proof.

If we again look at the situation in the rest frame we can write 
\begin{equation}W_i = \frac{1}{2} \epsilon_{i 0 jk}m M^{jk} = mS_i,
\label{eq:PL_restframe}
\end{equation}
where $S_i = \frac{1}{2} \epsilon_{ijk} M^{jk}$ is the {\bf spin operator}. \footnote{Observe that this discussion is problematic for massless particles. However, it is possible to find a similar relation for massless particles, when we chose a frame where the momentum of the particle is in a single direction.}
By showing that $WP=0$ we also have $W_0 = 0$ in this reference frame. This gives $W^2 = -\mathbf W^2 = -m^2\mathbf S^2$. Since the spin operator acts on a state with spin $s$ as $\mathbf S^2| s\rangle=s(s+1)| s\rangle$, we have that
\[W^2|m,s\rangle = -m^2 s(s+1)|m,s\rangle.\]
In addition this, also leads to the immediate conclusion that for each irrep with spin $s$ there are $2s+1$ {\it states} with spin component $s_3=0,\frac{1}{2},1,\frac{3}{2},\ldots$.

>>>>

So, what is the second Casimir of the Poincaré algebra?
\df{We define the {\bf Pauli-Ljubanski polarisation vector} by:
\begin{equation}
W_\mu \equiv \frac{1}{2} \epsilon_\mu{}_ \nu{}_\rho{}_\sigma P^\nu M^{\rho\sigma}.
\end{equation}
where $ \epsilon_\mu{}_ \nu{}_\rho{}_\sigma$ is the totally antisymmetric Levi-Civita tensor with $ \epsilon_{0123}=1$.}
We can show that this vector is translation invariant, {\it i.e.}\ that it commutes with the translation operator,
\begin{equation}
[P_\mu, W_\nu]=0.
\label{eq:PW_commutator}
\end{equation}
Then $W^2 = W_\mu W^\mu$ is a Casimir operator of the Poincaré algebra since we can show that
\begin{eqnarray}
\left[P_\mu, W^2\right] &=& 0,\\
\left[M_\mu{}_\nu, W^2\right] &=& 0.
\end{eqnarray}
Note that the second of these relationships is not trivial to demonstrate. See \cite{IntrSUSY2010} for a complete proof.

If we again look at the situation in the rest frame we can write 
\begin{equation}W_i = \frac{1}{2} \epsilon_{i 0 jk}m M^{jk} = mS_i,
\label{eq:PL_restframe}
\end{equation}
where $S_i = \frac{1}{2} \epsilon_{ijk} M^{jk}$ is the {\bf spin operator}. \footnote{Observe that this discussion is problematic for massless particles. However, it is possible to find a similar relation for massless particles, when we chose a frame where the momentum of the particle is in a single direction.}
By showing that $WP=0$ we also have $W_0 = 0$ in this reference frame. This gives $W^2 = -\mathbf W^2 = -m^2\mathbf S^2$. Since the spin operator acts on a state with spin $s$ as $\mathbf S^2| s\rangle=s(s+1)| s\rangle$, we have that
\[W^2|m,s\rangle = -m^2 s(s+1)|m,s\rangle.\]
In addition, this also leads to the immediate conclusion that for each irrep with spin $s$ there are $2s+1$ {\it states} with spin component $s_3=0,\frac{1}{2},1,\frac{3}{2},\ldots$.
